Ownership types in theory and practice (in Rust)

### 2022-11-27 paper.tex
Took some more notes
Wrote abstract

% - what are programming types in general?
% - why do we need ownership types, what do they do?

@@

\keywords{Rust, Ownership, Type, Safety}

@@

\subsection{History of Ownership Types}

@@

% - source: \cite{understanding-evolving-rust}
% - source: \cite{ownership-types-survey}

>>>>

% - why do we need ownership types, what do they do?

Object aliasing is the concept of accessing memory through different symbolic names in object-oriented programming languages.
Many progamming bugs are created through unintentional aliases, which are hard to detect and can lead to unexpected side-effects.
Ownership types are one solution that attempt to prevent many alias-related bugs.
There are different specific implementations of ownership types, but the premise is that 
The premise of ownership types is that not only the fields of an object are protected from external access, but also all (recursively defined) objects stored in those fields.
This is done by allowing objects to take ownership of other objects.
% TODO: add citation\cite{ownership-types-survey}

This paper introduces the different kinds of ownership types and explains how the modern programming language Rust uses ownership types to garantue memory safety.

@@

\keywords{Ownership, Type, Safety, Rust}

@@

\subsection{State of the art}

@@

% - source: \cite{understanding-evolving-rust}
% - source: \cite{ownership-types-survey}
% Notes from Ownership Types Survey\cite{ownership-types-survey}

% before: simply protect fields of an object from external access
% with OT: also protect objects stored in those fields by claiming ownership of and access to other objects
% problem OTs solve: object aliasing in OOP
% bugs due to unintentional aliases, difficult to find
% potential solutions: banning aliases, clearly advertising aliases, manage and control its effects
% page 16:
% 1998: flexible alias protection:
% limit visibility of changes to objects via aliases
% by limiting where alias could propagate
% by limiting the changes that could be observed through alias
% didn't feature "this"
% evolved into original OT:
% owners-as-dominators invariant:
% - heap divided into next regions (ownership contexts)
% - OC is set of objects
% - organised hierarchically: each owned object inside its owner, nesting relationship
% - each object has one single context
% - receives permissions to reference objects in other contexts through type

% core concept: object ownership: objects owner by other object or entities (global owners, stack-based owners)
% objects in same ownership context: siblings or peers

% control aliasing by removing aliasing
% unique references: only reference to object in system
% borrowing: 
% - reference to owned object temporarily passed to another object (for method call)
% - method over => all temporary references vanish/unusable

% control aliasing by controlling effects of aliasing
% immutability: state cannot change

% owners-as-dominators
% - program heap is tree structured, object is inside its owner
% - external access only through interface, i.e. owner
% - rules for a to validly access b: 1. a is owner of b; 2. a and b are siblings; 3. b is outside a
% - "a is inside b": "a owns b": a = b or a is owned by b, transitive
% - "a is outside b": converse of inside: "b is inside a"
% - advantages: simple, clear and strong garantee, easy to reason about properties of code
% - disadvantages: no aliasing => no common idioms; programming more difficult

% owners-as-modifiers
% - weaker form of owners-as-dominators, allows read-only references
% - read-only reference: only used to read fields, call pure methods (may not modify any existing object, including receiver)
% - aliasing unrestricted, but only owner can modify
% - usecase: requirements from verification of object-oriented programs
% - rules for a to validly reference b through reference r: either
% 1. a is owner of b
% 2. a and b are siblings
% 3. b is outside a
% 4. r is read-only reference and only pure methods can be called

% ownership domains:

% owners-as-ombudsmen:
% - problem: strong topological requirement: tree structure: some idioms don't fit in tree structure
% - allow multiple objects to define a shared aggregate owner
% - rules for a to validly reference b through reference r: either
% 1. a is owner of b
% 2. a and b are siblings
% 3. b is outside a
% 4. a is owned by aggregate owner b

% QUESTIONS:
% - difference between aliases and references
% - difference between "a is owner of b" and "b is outside a"
